---
title: "hw03-lindsay-dahlen.Rmd"
author: "Lindsay Dahlen"
date: "October 15, 2017"
output: github_document
---
```{r}
# Ranking by Salary
library(dplyr)
library(readr)
library(ggplot2)
setwd("~/Stat 133/stat133-hws-fall17/hw03/data")
teams <- read_csv("nba2017-teams.csv")
summary(teams)
teams_salary <- arrange(teams, desc(salary))
head(teams_salary)

salary_mean <- mean(teams_salary$salary)
```  
## ggplot(teams_salary, aes(x = team, y = salary)) + geom_bar() + coord_flip() +  geom_hline(yintercept = salary_mean) + xlab(Team) + ylab(Salary(in millions)) +  ggtitle("NBA Teams ranked by Total Points", subtitle = NULL) # Help



```{r}
# Ranking by Points
teams_points <- arrange(teams, desc(points))
head(teams_points)

points_mean <- mean(teams_points$points)

```
## ggplot(teams_points, aes(x = teams, y = points)) +
  geom_bar() +
  coord_flip() +
  geom_hline(yintercept = points_mean, col = "red") +
  xlab("Total Points") +
  ylab("Team") + 
  ggtitle("NBA Teams ranked by Total Points", subtitle = NULL)
# Help

```{r}
# Ranking by Efficiency
teams_efficiency <- arrange(teams, desc(efficiency))
head(teams_efficiency)

efficiency_mean <- mean(teams_efficiency$efficiency)

```
## ggplot(teams_efficiency, aes(x = teams, y = efficiency)) +
  geom_bar() +
  coord_flip() +
  geom_hline(yintercept = efficiency_mean, col = "red") +
  xlab("Total Efficiency") +
  ylab("Team") + 
  ggtitle("NBA Teams ranked by Total Efficiency", subtitle = NULL)

# Help


The tables show that Cleveland is the top team when ranked by salary and efficiency, despite not scoring the most points. It also shows that total salary is not linked to total points or total efficiency.

```{r}
# Principle Components Analysis
pca_points3 <- prcomp(teams$points3, scale. = TRUE)
pca_points2 <- prcomp(teams$points2, scale. = TRUE)
pca_free_throws <- prcomp(teams$free_throws, scale. = TRUE)
pca_off_rebounds <- prcomp(teams$off_rebounds, scale. = TRUE)
pca_def_rebounds <- prcomp(teams$def_rebounds, scale. = TRUE)
pca_assists <- prcomp(teams$assists, scale. = TRUE)
pca_steals <- prcomp(teams$steals, scale. = TRUE)
pca_blocks <- prcomp(teams$blocks, scale. = TRUE)
pca_turnovers <- prcomp(teams$turnovers, scale. = TRUE)
pca_fouls <- prcomp(teams$fouls, scale. = TRUE)

eigs <- data.frame(eigenvalue = pca_points3$sdev^2, proportion = round(100*pca_points3$sdev^2 / sum(pca_points3$sdev^2), 3))
eigs

```