---
title: "hw01-lindsay-dahlen.Rmd"
author: "Lindsay Dahlen"
date: "September 20, 2017"
output: github_document
---
#HW 01
```{r}
load("data/nba2017-salary-points.RData")
ls()
```

## 1)
```{r}
salarymil <- salary/1000000
salarymiltwo <- round(salarymil, digits = 2)
salarymiltwo
```
```{r}
# Experience 
experienceN <- replace(experience, experience == "R", 0)
experience_integer <- as.integer(experienceN)
class(experience_integer)
experienceN
```
```{r}
# Position
fposition <- factor(position)
fposition
levels(fposition) <- c('center', 'power_fwd', 'point_guard', 'small_fwd', 'shoot_guard')
fposition
```
```{r}
# Calculate frequencies
table(fposition)
```
## 2)
```{r}
# Plot
plot(points, salary/1000000, pch = 15, col = "chartreuse", cex = 1, xlab = "points", ylab = "salary (in millions of dollars", main = "Points vs. Salary")
```
 ### This scatterplot indicates that more points scored correlates with a higher salary, although there are exceptions to this. For example, some players that score many points have similar salaries to players that score fewer points, and vice versa.
 
 ## 3)
 
```{r}
# Finding correlation
n <- 441      # There are 441 players
n
# Mean
mean_points <- sum(points)/441
mean_points

mean_salarymiltwo <- sum(salarymiltwo)/441
mean_salarymiltwo
```
```{r}
# var(X)
var_points1 <- (points - mean_points)
var_points2 <- sum(var_points1^2)
var_points <- (1/440)*var_points2
var_points

# var(Y)
var_salarymiltwo1 <- (salarymiltwo - mean_salarymiltwo)
var_salarymiltwo2 <- sum(var_salarymiltwo1^2)
var_salarymiltwo <- (1/440)*var_salarymiltwo2
var_salarymiltwo
```
```{r}
# Standard deviation of X
sd_points <- sqrt(var_points)
sd_points

# Standard deviation of Y
sd_salarymiltwo <- sqrt(var_salarymiltwo)
sd_salarymiltwo
```
```{r}
# Covariance between X and Y
cov_p_s_1 <- (points - mean_points)*(salarymiltwo - mean_salarymiltwo)
cov_p_s_2 <- sum(cov_p_s_1)
cov_p_s <- (1/440)*cov_p_s_2
cov_p_s
```
```{r}
# Correlation
cor_points_salarymiltwo <- (cov_p_s)/(sd_points * sd_salarymiltwo)
cor_points_salarymiltwo

cor(points, salarymiltwo, use = "everything", method = c("pearson", "kendall", "spearman")) # Used cor() to confirm correlation I found manually.
```
\overline{x} = $$\sum_{i=1}^{n} X_i$$

## 4)
```{r}
# Estimated slope of the regression line
slope1 <- cor_points_salarymiltwo * (sd_salarymiltwo/sd_points)
slope1
```
```{r}
# Estimated intercept of the regression line
intercept0 <- mean_salarymiltwo - (slope1*mean_points)
intercept0
```
```{r}
# Y-hat, or predicted values of Y
y_hat <- intercept0 + (slope1*points)
y_hat
summary(y_hat)
x_hat <- intercept0 + (slope1*salarymiltwo)
summary(x_hat)
```
- The regression equation is $ yhat = b_{0}+b_{1}X$ 
- The slope coefficient $b_{1}$ represents how many points a player needs to score in order to increase their salary.
- The intercept term $b_{0}$ represents the minimum points a player can score in order to receive a salary.
- The predicted salary of a player that scores:
-- 0 points is $1.509077 million.
-- 100 points is $2.3657567 million.
-- 500 points is $5.7887835 million.
-- 1000 points is $10.067567 million.
-- 2000 points is $18.625134 million.
-----------------------------------------
## 5)
```{r}
plot(points, salary/1000000, pch = 15, col = "chartreuse", cex = 1, xlab = "points", ylab = "salary (in millions of dollars", main = "Points vs. Salary")
abline(a = intercept0, b = slope1, col = "midnightblue", lwd = 2)
lines(lowess(points, salarymiltwo), col = "red", lwd = 2)
text(2500, 30, "lowess", col = "red", cex = 1)
text(2450, 20, "regression", col = "midnightblue", cex = 1)
```

------------------------------------------
## 6)
```{r}
# Regression Residuals
# Residaul of Points
residuals_points <- points - y_hat
summary(residuals_points)

# Residual of Salary
residuals_salary <- salarymiltwo - x_hat
summary(residuals_salary)

# Residual Sum of Squares for Points
RSS_points <-sum(residuals_points^2)
RSS_points

# Residual Sum of Squares for Salary in millions
RSS_salarymiltwo <- sum(residuals_salary^2)
RSS_salarymiltwo

# Total Sum of Squares 
TSS_salary <- sum((salarymiltwo - mean_salarymiltwo)^2)
TSS_points <- sum((points - mean_points)^2)
TSS_salary
TSS_points

# The coefficient of determination for Points
R2_points <- 1-(RSS_points/TSS_points)
R2_points

# The coefficient of determination for Salary in millionis
R2_salary <- 1-(RSS_salarymiltwo/TSS_salary)
R2_salary
```
-------------------------------
## 7)
```{r}
# Scatterplot of Years of Experience and Salary
plot(experience_integer, salarymiltwo, type = "p", main = "Scatterplot of Salary and Years of Experience", xlab = "Experience in Years", ylab = "Salary in Millions of Dollars", ann = par("ann"), cex = 1, pch = 21)
lines(loess.smooth(experience_integer, salarymiltwo, span = 2/3, degree = 1,
    family = c("symmetric", "gaussian"), evaluation = 50), col = "red", lwd = 2)
```
```{r}
# Scatterplot of Points, Experience, and Salary
library(scatterplot3d)
scatterplot3d(points, experience_integer, salarymiltwo, color = par("col"), pch = 1, main = "Points, Experience, and Salary Scatterplot", xlab = "Points", ylab = "Experience in years", zlab = "Salary in Millions of Dollars", scale.y = 1, col.grid = "grey")
```
```{r}
# Conditional boxplot of Salary in terms of Position
boxplot(salarymiltwo ~ position, range = 1, varwidth = TRUE, outline = TRUE, col = "papayawhip", xlab = "Position", ylab = "Salary in Millions of Dollars")
```
- The first plot seems to indicate that years of experience does not have anything to do with the salary. The second plot shows that in general, more years of experience does not equate to a higher salary, but that more points scored does. The last plot shows that in general, all positions are paid around the same salary.
- Based on the scatterplots, experience seems to be somewhat related with salary, though salary is more related to points than experience.
- Based on the boxplot, position does not seem to be related with salary.

-------------------------------
## 8)
- Things that were hard were formating the in-text code, looking up functions I didn't know, and formating the plots.
- One thing that was easy-ish was creating new chunks without having to press the 'insert' button.
- This class is my first exposure to Git. I don't think I like it, mainly because I'm unsure about what I'm doing.
- I did need help to complete this assignment. I got assistance from my friend Hannah S. who is also taking the class, and my friend Sruthi.
- This homework took me over 8 hours to complete, spread over 3 days.
- The most time consuming part was trying to look up functions, especially for graphing data, and trying to make it work. It was also tedious to write out each equation in parts.
- I still do not understand how to write in-line equations.
- The most frustrating thing about this homework was how helpless I felt. I dislike having to look up everything myself, especially since I'm not sure if I'm doing it correctly.
- I'm very proud of my plots, especially with changing the colors/labels/shapes/etc.
```

