---
title: "post02-lindsay-dahlen"
author: "Lindsay Dahlen"
date: "December 3, 2017"
output: html_document
---
### ggvis: Visualizing data and its properties

1. ggvis

In this post I will talk about the data visualization package ggvis. We were required to use ggvis in our interactive shiny app as part of homework 4. I hope to expand on what exactly ggvis is and various ways you can use it, particularly in comparison to the package ggplot2. I will not be using a shiny app in this post, but will be using ggvis in Rstudio.

2. Introduction
ggvis is a package that combines the data computing power of R and the graphic rendering power of web browsers, using "Vega". (Vega is a visualization grammar, which I won't be discussing in this post). Rstudio can display ggvis because its a web browser. The graphics will display in the viewer panel.

3. Set-up
In the following examples I will be using the Iris dataset, which was created by R. A. Fisher using data collected by E. Anderson. It "iris is a data frame with 150 cases (rows) and 5 variables (columns) named *Sepal.Length*, *Sepal.Width*, *Petal.Length*, *Petal.Width*, and *Species*."
```{r}
# Summary of iris dataset
summary(iris)
head(iris)
```
I installed ggvis as follows:
```{r}
install.packages("ggvis")
# intsall latest version
devtools::install_github("hadley/lazyeval", build_vignettes = FALSE)
devtools::install_github("hadley/dplyr", build_vignettes = FALSE)
devtools::install_github("rstudio/ggvis", build_vignettes = FALSE)
library(ggvis)
```
I also loaded the *dplyr* package, which is necessary to write for ggvis. Additionally, I loaded the *ggplot2* package to show comparisons between ggplot and ggvis.
```{r}
library(dplyr)
library(ggplot2)
```


4. The Basics of ggvis
The components of ggvis include **data**, or the data frame; **layer**, or the layers of the plot componenets; **mappings**, which map between variables and aesthetics; and **~**, which are equivalent to aes mappings in ggplot2.

Unlike ggplot2, ggvis uses the pipeline operator to 

8. ggplot2 vs. ggvis

Using ggvis is very similar to using ggplot2. There are some notable distinctions, which I will summarize as follows.
 
- Faceting is not supported in ggvis. Even when it it, it does not resember faceting in ggplot2.
- 

9. Conclusion

10. References
- https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/iris.html
- https://ggvis.rstudio.com/
- https://cran.r-project.org/web/packages/ggvis/README.html
- http://ggvis.rstudio.com/ggplot2.html
- http://jimhester.github.io/ggplot2ToGgvis/
-

